---
title: "R Notebook"
output: html_notebook
---
# Load in the libraries
```{r}
library(tidyverse)
library(here)
library(stringr)
```

#Read in clean csv
```{r}
candy <- read_csv(here("clean_data/candy_clean.csv"))
```


# 1. What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
```{r}
candy %>%
  filter(!is.na(rating)) %>% 
  summarise(count_rating = n())
```
The above code chunk was attempting to find the total number of candy ratings
It was decided that all NA ratings should be dropped via a `filter`
The total count was then calculated by using a `summarise`



#2. What was the average age of people who are going out trick or treating?
```{r}
candy %>% 
  filter(trick_or_treating == "Yes") %>% 
  summarise(avg_age_tt = floor(mean(age)))
```
The aim here was to find the average age of people who went trick or treating
The data was filtered to only return people that went trick or treating
The average age was then calculated by using the `summarise` function and finding the `mean` age
It was decided to `floor` this avg_age value to return the nearest age


# 3. What was the average age of people who are not going trick or treating?
```{r}
candy %>% 
  filter(trick_or_treating == "No") %>% 
  summarise(avg_age_no_tt = floor(mean(age)))
```
The aim here was to find the average age of people who didn't go trick or treating
The data was filtered to only return people that didn't go trick or treating
The average age was then calculated by using the `summarise` function and finding the `mean` age
It was decided to `floor` this avg_age value to return the nearest age


# 4. For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
candy %>% 
  filter(rating == "JOY") %>% 
  group_by(candies, rating) %>% 
  summarise(joy_rating_per_candies = n()) %>% 
  arrange(desc(joy_rating_per_candies)) %>% 
  head(1) %>% 
  ungroup()

candy %>% 
  filter(rating == "DESPAIR") %>% 
  group_by(candies, rating) %>% 
  summarise(joy_rating_per_candies = n()) %>% 
  arrange(desc(joy_rating_per_candies)) %>% 
  head(1) %>% 
  ungroup()

candy %>% 
  filter(rating == "MEH") %>% 
  group_by(candies, rating) %>% 
  summarise(joy_rating_per_candies = n()) %>% 
  arrange(desc(joy_rating_per_candies)) %>% 
  head(1) %>% 
  ungroup()
```
The aim here was to find the top rated candy bars for each of the 3 rating categories
3 separate code chunks were written to return 3 separate answers
Each code was filtered to match the rating category
The data was then grouped by the type of candy, followed by the rating category
These ratings were then counted using `summarise` 
The data was then arranged in descending order to find the candy with the most ratings for each category
The top result for each was taken by using the `head` function


# 5. How many people rated Starburst as despair?
```{r}
candy %>% 
  filter(candies == "starburst") %>% 
  filter(rating == "DESPAIR") %>% 
  summarise(despair_rating_count_starburst = n())
```
This question asked to find how many people rated Starburst as DESPAIR
The data was filtered by candy and rating to only return the relevant results
The value was then counted by using the `summarise` verb


# For the next three questions, count despair as -1, joy as +1, and meh as 0.
```{r}
candy_count <- candy %>% 
  mutate(rating = str_replace_all(rating, "JOY", "1"),
         rating = str_replace_all(rating, "DESPAIR", "-1"),
         rating = str_replace_all(rating, "MEH", "0"))

candy_count <- candy_count %>% 
  mutate(rating = as.numeric(rating))
```
For the next series of questions, it was decided to recode the rating system to numeric values
These values could then be summed to return a numeric value


# 6. What was the most popular candy bar by this rating system for each gender in the dataset ?
```{r}
candy_count %>%
  filter(!is.na(rating)) %>%
  filter(gender == "Male") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_by_male = sum(rating)) %>% 
  arrange(desc(pop_candy_by_male)) %>% 
  head(1) %>% 
  ungroup()

candy_count %>%
  filter(!is.na(rating)) %>%
  filter(gender == "Female") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_by_female = sum(rating)) %>% 
  arrange(desc(pop_candy_by_female)) %>% 
  head(1) %>% 
  ungroup()

candy_count %>%
  filter(!is.na(rating)) %>%
  filter(gender == "I'd rather not say") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_by_other = sum(rating)) %>% 
  arrange(desc(pop_candy_by_other)) %>% 
  head(1) %>% 
  ungroup()

candy_count %>%
  filter(!is.na(rating)) %>%
  filter(is.na(gender)) %>% 
  group_by(candies) %>% 
  summarise(pop_candy_by_NA = sum(rating)) %>% 
  arrange(desc(pop_candy_by_NA)) %>% 
  head(1) %>% 
  ungroup()
```
There were 4 gender categories that were filtered to find the most popular candy bar in each
The data was first of all filtered to only return candies with a rating
The data was then filtered by gender
Candies were grouped together 
`Summarise` was then used to sum the ratings for each candy (1 for JOY, 0 for MEH, -1 for DESPAIR)
The results were then arranged in descending order before taking the top result for each


# 7. What was the most popular candy bar in each year?
```{r}
candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(year == 2015) %>% 
  group_by(candies) %>% 
  summarise(pop_candy_2015 = sum(rating)) %>% 
  arrange(desc(pop_candy_2015)) %>% 
  head(1) %>% 
  ungroup()


candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(year == 2016) %>% 
  group_by(candies) %>% 
  summarise(pop_candy_2016 = sum(rating)) %>% 
  arrange(desc(pop_candy_2016)) %>% 
  head(1) %>% 
  ungroup()


candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(year == 2017) %>% 
  group_by(candies) %>% 
  summarise(pop_candy_2017 = sum(rating)) %>% 
  arrange(desc(pop_candy_2017)) %>% 
  head(1) %>% 
  ungroup()
```
The years were filtered to find the most popular candy bar in each
The data was first of all filtered to only return candies with a rating
The data was then filtered by years
Candies were grouped together 
`Summarise` was then used to sum the ratings for each candy (1 for JOY, 0 for MEH, -1 for DESPAIR)
The results were then arranged in descending order before taking the top result for each


# 8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
```{r}
candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(country == "USA") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_USA = sum(rating)) %>% 
  arrange(desc(pop_candy_USA)) %>% 
  head(1) %>% 
  ungroup()


candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(country == "Canada") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_Canada = sum(rating)) %>% 
  arrange(desc(pop_candy_Canada)) %>% 
  head(1) %>% 
  ungroup()


candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(country == "UK") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_UK = sum(rating)) %>% 
  arrange(desc(pop_candy_UK)) %>% 
  head(1) %>% 
  ungroup()


candy_count %>% 
  filter(!is.na(rating)) %>%
  filter(!is.na(country)) %>% 
  filter(!(country %in% c("USA", "UK", "Canada"))) %>% 
  group_by(candies) %>% 
  summarise(pop_candy_others = sum(rating)) %>% 
  arrange(desc(pop_candy_others)) %>% 
  head(1) %>% 
  ungroup()
```
There were 4 categories of countries that were filtered to find the most popular candy bar in each
The data was first of all filtered to only return candies with a rating
The data was then filtered by country
Candies were grouped together 
`Summarise` was then used to sum the ratings for each candy (1 for JOY, 0 for MEH, -1 for DESPAIR)
The results were then arranged in descending order before taking the top result for each
For the last group, it was decided to find all the countries, apart from USA, UK & Canada
