---
title: "Task_04_Halloween_Candy"
author: "Lloyd Lombardi"
date: "`r Sys.Date()`"
output: html_document
---
# Task 4 - Halloween Candy

```{r, echo=FALSE}
library(meme)
u <- "https://www.care.com/c/wp-content/uploads/sites/2/2021/10/Screen-Shot-2021-09-30-at-8.36.29-AM-1024x993.png"
meme(u)
```

## Introduction
The purpose of this Markdown file is to give an overview of Task 4 of the **Dirty Data Project**.

The task concerned 3 `.xlsx` files which contained data about **Halloween Candy** over three years:
**2015, 2016 & 2017**.

The data was in an **untidy** format, and therefore needed to be cleaned before
the following questions could be answered:

1. What is the total number of candy ratings given across the three years?
2. What was the average age of people who are going out trick or treating?
3. What was the average age of people who are not going trick or treating?
4. For each of joy, despair and meh, which candy bar received the most of these ratings?
5. How many people rated Starburst as despair?
6. What was the most popular candy bar by this rating system for each gender in the dataset ?
7. What was the most popular candy bar in each year?
8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?


The data had 3 separate `.xlsx` files which contained similar information, but a **lot** of cleaning was required before any analysis could be conducted.


In the rest of this document, the following information will be presented:

- A list of assumptions made
- Steps taken to clean the data
- Answers to the analysis questions


## Assumptions Made

- Only seven columns were required to **fully answer the analysis questions**
    1. candies
    2. year
    3. rating
    4. age
    5. gender
    6. country
    7. trick or treating
- Columns could be renamed to match candies from other data sets
- Data should be **pivoted longer** to make analysis easier
- All candy was kept in from all tables
- The age column could be filtered to only return ages between 0 and 100
- The 'extra' columns towards the end of each data sets that asked users to input other candies to give them **JOY** or **DESPAIR** could be dropped as this would have no affect on the final analysis questions
- The 2015 data set could have **gender** and **country** columns added to them, with **NA** values for all
- Any users that input a **US State** as their country had this changed to **USA**
- Any users that input a **Home Nation** as their country had this changed to **UK**
- Any users that input an unrecognisable country had their input **dropped**
- Trick or treating column could be converted to a logical column


## Steps Taken to Clean the Data

The tables were cleaned separately before being joined together

### 2015 Cleaninig

- The first step was to rename some columns to match the 2016 & 2017 candies
```{r, eval=FALSE}
candy_2015 <- candy_2015 %>% 
  rename(boxo_raisins = box_o_raisins,
         hersheys_kisses = hershey_s_kissables,
         licorice_yes_black = licorice)
```

- The age columns were then tidied and wrangled
- Any age 100 or over was removed
- An age equivalent to pi was removed
- Various ages with other syntax were tidied to only return a numerical value
```{r, eval=FALSE}
candy_2015_ages <- candy_2015 %>% 
  mutate(how_old_are_you = str_remove_all(how_old_are_you, "[0-9][0-9][0-9]"),
         how_old_are_you = str_remove_all(how_old_are_you, "3.14%"),
         how_old_are_you = str_replace(how_old_are_you, ">39", "39"),
         how_old_are_you = str_replace(how_old_are_you, "７１＋", "71"),
         how_old_are_you = str_replace(how_old_are_you, "Good Lord!  I'm 43!", "43"))
```

- The candy columns were pivoted longer
- A range of columns were selected for the updated data set
- Two further columns were named as they fell outside of this range
```{r, eval=FALSE}
candy_2015_longer <- candy_2015_ages %>% 
  pivot_longer(cols = c(butterfinger:york_peppermint_patties, 
                        sea_salt_flavored_stuff_probably_chocolate_since_this_is_the_it_flavor_of_the_year, 
                        necco_wafers),
               names_to = "candies",
               values_to = "rating")
```

- A **year** column was added with 2015 for all values
- The candies, rating and year columns were moved to the front of the data set
- 5 of the 7 columns stated in the assumptions were selected
- 2 columns were renamed to **age** and **trick_or_treating**
```{r, eval=FALSE}
candy_2015_clean <- candy_2015_longer %>% 
  mutate(year = 2015) %>% 
  relocate(c(candies:year), .after = timestamp) %>% 
  select(c(candies:are_you_going_actually_going_trick_or_treating_yourself)) %>% 
  rename(age = how_old_are_you,
         trick_or_treating = are_you_going_actually_going_trick_or_treating_yourself)
```

- The age column was changed to a **numeric** type
- One or more number characters were extracted before being converted
```{r, eval=FALSE}
candy_2015_clean <- candy_2015_clean %>% 
  mutate(age = str_extract(age, "^[0-9]+")) %>% 
  mutate(age = as.numeric(age))
```

- The ages were then filtered to meet the above assumption
```{r, eval=FALSE}
candy_2015_clean <- candy_2015_clean %>% 
  filter(age > 0,
         age < 100)
```

- A country and gender were added to complete the 7 required columns
- These were populated with NA values
```{r, eval=FALSE}
candy_2015_clean <- candy_2015_clean %>% 
  mutate(country = "",
         country = na_if(country, ""),
         gender = "",
         gender = na_if(gender, ""))
```


### 2016 Cleaning

- The age column was the first thing to be tidied
- Any age 100 or over was removed
- Various ages with other syntax were tidied to only return a numerical value
```{r, eval=FALSE}
candy_2016_ages <- candy_2016 %>% 
  mutate(how_old_are_you = str_remove_all(how_old_are_you, "[0-9][0-9][0-9]"),
         how_old_are_you = str_replace(how_old_are_you, "49 11/12ths", "49"),
         how_old_are_you = str_replace(how_old_are_you, "Fifty.  Nine.  Ish.", "59"))
```

- The candy columns were pivoted longer
- A range of columns were selected for the updated data set
```{r, eval=FALSE}
candy_2016_longer <- candy_2016_ages %>% 
  pivot_longer(cols = c(x100_grand_bar:york_peppermint_patties),
               names_to = "candies",
               values_to = "rating")
```

- Columns were wrangled
- A **year** column was added with 2016 for all values
- The candies, rating and year columns were moved to the front of the data set
- 7 columns stated in the assumptions were selected
- 4 columns were renamed to **age**, **trick_or_treating**, **gender** and **country** to match the 2015 table
```{r, eval=FALSE}
candy_2016_clean <- candy_2016_longer %>% 
  mutate(year = 2016) %>% 
  relocate(c(candies:year), .after = timestamp) %>% 
  select(c(candies:which_country_do_you_live_in)) %>% 
  rename(age = how_old_are_you,
         trick_or_treating = are_you_going_actually_going_trick_or_treating_yourself,
         gender = your_gender,
         country = which_country_do_you_live_in)
```

- Some columns were moved again to match 2015 exactly
```{r, eval=FALSE}
candy_2016_clean <- candy_2016_clean %>% 
  relocate(age, .after = year)

candy_2016_clean <- candy_2016_clean %>% 
  relocate(gender, .after = country)
```

- The age column was changed to a **numeric** type
- One or more number characters were extracted before being converted
```{r, eval=FALSE}
candy_2016_clean <- candy_2016_clean %>% 
  mutate(age = str_extract(age, "^[0-9]+")) %>% 
  mutate(age = as.numeric(age))
```

- The ages were filtered to match above
```{r, eval=FALSE}
candy_2016_clean <- candy_2016_clean %>% 
  filter(age > 0,
         age < 100)
```


### 2017 Cleaning

- Remove the 'q1' style prefix to the column names
```{r, eval=FALSE}
candy_2017 <- candy_2017 %>% 
  rename_with(~ gsub("^q[0-9]_", "", .x))
```

- Add an 'x' onto the name of a candy to match the 2015 & 2016 data sets
```{r, eval=FALSE}
candy_2017 <- candy_2017 %>% 
  rename_with(~str_replace(., "^", "x"), .cols = starts_with("100"))
```

- Rename a candy to match the 2015 & 2016 data sets
```{r, eval=FALSE}
candy_2017 <- candy_2017 %>% 
  rename(mary_janes = anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes)
```

- The age column was the first thing to be tidied
- Any age 100 or over was removed
- Various ages with other syntax were tidied to only return a numerical value
```{r, eval=FALSE}
candy_2017_ages <- candy_2017 %>% 
  mutate(age = str_remove_all(age, "[0-9][0-9][0-9]"),
         age = str_replace(age, "45-55", "50"),
         age = str_replace(age, "24-50", "37"),
         age = str_replace(age, "sixty-nine", "69"))
```

- The candy columns were pivoted longer
- A range of columns were selected for the updated data set
```{r, eval=FALSE}
candy_2017_longer <- candy_2017_ages %>% 
  pivot_longer(cols = c(x100_grand_bar:york_peppermint_patties),
               names_to = "candies",
               values_to = "rating")
```

- Column wrangling
- A **year** column was added with 2016 for all values
- The candies, rating and year columns were moved to the front of the data set
- 7 columns stated in the assumptions were selected
- 1 column1 was renamed to **trick_or_treating**to match the 2015 & 2016 table
```{r, eval=FALSE}
candy_2017_clean <- candy_2017_longer %>% 
  mutate(year = 2017) %>% 
  relocate(c(candies:year), .after = internal_id) %>% 
  select(c(candies:country)) %>% 
  rename(trick_or_treating = going_out)
```

- Some columns were moved again to match 2015 & 2016 exactly
```{r, eval=FALSE}
candy_2017_clean <- candy_2017_clean %>% 
  relocate(age, .after = year)

candy_2017_clean <- candy_2017_clean %>% 
  relocate(gender, .after = country)

```

- The age column was changed to a **numeric** type
- One or more number characters were extracted before being converted
```{r, eval=FALSE}
candy_2017_clean <- candy_2017_clean %>% 
  mutate(age = str_extract(age, "^[0-9]+")) %>% 
  mutate(age = as.numeric(age))
```

- The ages were filtered to match above
```{r, eval=FALSE}
candy_2017_clean <- candy_2017_clean %>% 
  filter(age > 0,
         age < 100)
```


### Joining Tables

It was decided that `bind_rows` was the best option to join the tables as they all had identical column names
```{r, eval=FALSE}
candy <- bind_rows(candy_2015_clean, candy_2016_clean, candy_2017_clean)
```


### Cleaning the Country Column

The **country** column was particularly messy, and it was decided that it would be more efficient to clean this after the tables were joined

There were **115** `distinct` countries in the raw data, and when cleaned there were **39**.

- The countries were changed `to_title` to merge some values together
```{r, eval=FALSE}
candy <- candy %>% 
  mutate(country = str_to_title(country))
```

- Vectors were created to capture some information to be converted into a country
```{r, eval=FALSE}
states <- c("North Carolina", "Pittsburgh", "New York", "California", "New Jersey", "Alaska")
uk <- c("England", "Scotland", "Endland")
```

- Regex was used to convert variations of USA, UK, Canada, Spain and Netherlands to one input per country
- `case_when` was used to detect strings in this column and then change that to a stated string
- A default value was then stated for all other countries
```{r, eval=FALSE}
candy_country_sorted <- candy %>% 
  mutate(country = case_when(
    str_detect(country,"((?i)[a-z]* *[:punct:])*((?i)^u+ *[:punct:]*(?i)[ns]+ *[:punct:]*(?i)s* *[:punct:]*(?i)a* *[:punct:]*)+((?i)[a-z]* *[:punct:])*") ~ "USA",
    str_detect(country, "(?i)a+mer") ~ "USA",
    str_detect(country, "(?i)t+[a-z]* *(?i)u+[a-z]* *(?i)s+[a-z]* *(?i)o*[a-z]* *(?i)a*[a-z]*") ~ "USA",
    str_detect(country, " +U+[sa]*") ~ "USA",
    str_detect(country, "T+[a-z]* Y+[a-z]* E+[a-z]* O+[a-z]* A+[a-z]*") ~ "USA",
    str_detect(country, "[:punct:]*M+[ue]+[r]+") ~ "USA",
    country %in% states ~ "USA",
    str_detect(country, "^(?i)u+[:punct:]*(?i)k+") ~ "UK",
    country %in% uk ~ "UK",
    str_detect(country, "(^(?i)can)+[a-z]*[:punct:]*") ~ "Canada",
    str_detect(country, "ñ+") ~ "Spain",
    str_detect(country, "The N+") ~ "Netherlands",
    TRUE ~ country
  ))
```

- Another vector was created to capture 'other' countries that had no real value
```{r, eval=FALSE}
country_outliers = c("A tropical island south of the equator", "Ud", "Atlantis",
                     "Canae", "cascadia ", "Cascadia", "Denial", "Earth", "1", 
                     "god's country", "I Don't Know Anymore", "insanity lately", 
                     "There Isn't One For Old Men", "Soviet Canuckistan", "Narnia", "Neverland",
                     "one of the best ones", "See above", "Somewhere", "subscribe to dm4uz3 on youtube",
                     "The Republic Of Cascadia", "this one", "Eua", "Europe")
```

- The above 'other' countries were then converted to **NA** values as they were not useful for analysis
```{r, eval=FALSE}
candy_country_sorted <- candy_country_sorted %>% 
  mutate(country = if_else(country %in% country_outliers, NA_character_, country))
```


## Assertive Programming and Unit Testing

It was decided to introduce some **Assertive Programming** and **Unit Testing** to the cleaning script

The scope and span of these are admittedly quite narrow and not that useful, but they were left in to showcase their use

- The **trick_or_treating** column was converted to a logical type for the Unit Test
```{r, eval=FALSE}
candy_country_sorted <- candy_country_sorted %>% 
  mutate(trick_or_treating = recode(trick_or_treating,
                                    "Yes" = "TRUE",
                                    "No" = "FALSE"),
         trick_or_treating = as.logical(trick_or_treating))
```

- A unit test was then performed to check that numerical or character inputs would return a test passed
```{r, eval=FALSE}
test_that("Expect error if trick_or_treating is not logical", {
  expect_error(trick_or_treating(10))
  expect_error(trick_or_treating("Yes"))
})
```

- Assertive programming was used to verify that the ages were within the stated limits and the years were only 2015, 2016 or 2017
```{r, eval=FALSE}
candy_clean %>% 
  verify(age > 0 & age < 100) %>% 
  verify(year == 2015 | 2016 | 2017)
```


## Answers to the Analysis Questions

##### 1. What is the total number of candy ratings given across the three years?
```{r, eval=FALSE}
count_rating = 738763 
```

##### 2. What was the average age of people who are going out trick or treating?
```{r, eval=FALSE}
avg_age_tt = 35
```

##### 3. What was the average age of people who are not going trick or treating?
```{r, eval=FALSE}
avg_age_no_tt = 39
```

##### 4. For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r, eval=FALSE}
JOY:
          any_full_sized_candy_bar - 7268

DESPAIR:
          broken_glow_stick - 7561
  
MEH:
          lollipops - 1512
```

##### 5. How many people rated Starburst as despair?
```{r, eval=FALSE}
despair_rating_count_starburst - 1873
```

##### 6. What was the most popular candy bar by this rating system for each gender in the dataset ?
```{r, eval=FALSE}
Male:
          any_full_sized_candy_bar - 1537
  
Female:
          any_full_sized_candy_bar - 856
  
Other:
          any_full_sized_candy_bar - 51
  
NA: 
          any_full_sized_candy_bar - 4412
```

##### 7. What was the most popular candy bar in each year?
```{r, eval=FALSE}
2015:
          any_full_sized_candy_bar - 4405
  
2016:
          any_full_sized_candy_bar - 987
  
2017:
          any_full_sized_candy_bar - 1499
```

##### 8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
```{r, eval=FALSE}
USA: 
          any_full_sized_candy_bar - 2139

UK:
          cash_or_other_forms_of_legal_tender - 25

Canada:
          any_full_sized_candy_bar - 248

Others:
          any_full_sized_candy_bar - 65
```

### Check out the **Github Repo** [here](https://github.com/lloydlombardi/dirty_data_codeclan_project_lloyd_lombardi/tree/main/task_04_halloween_candy_data)