---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(here)
library(stringr)
```



```{r}
candy <- read_csv(here("clean_data/candy_clean.csv"))
```


# 1. What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
```{r}
candy %>%
  filter(!is.na(rating)) %>% 
  summarise(count_rating = n())
```

#2. What was the average age of people who are going out trick or treating?
```{r}
candy %>% 
  filter(trick_or_treating == "Yes") %>% 
  summarise(avg_age_tt = floor(mean(age)))
```

# 3. What was the average age of people who are not going trick or treating?
```{r}
candy %>% 
  filter(trick_or_treating == "No") %>% 
  summarise(avg_age_no_tt = floor(mean(age)))
```


# 4. For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
candy %>% 
  filter(rating == "JOY") %>% 
  group_by(candies, rating) %>% 
  summarise(joy_rating_per_candies = n()) %>% 
  arrange(desc(joy_rating_per_candies)) %>% 
  head(1) %>% 
  ungroup()

candy %>% 
  filter(rating == "DESPAIR") %>% 
  group_by(candies, rating) %>% 
  summarise(joy_rating_per_candies = n()) %>% 
  arrange(desc(joy_rating_per_candies)) %>% 
  head(1) %>% 
  ungroup()

candy %>% 
  filter(rating == "MEH") %>% 
  group_by(candies, rating) %>% 
  summarise(joy_rating_per_candies = n()) %>% 
  arrange(desc(joy_rating_per_candies)) %>% 
  head(1) %>% 
  ungroup()
```


# 5. How many people rated Starburst as despair?
```{r}
candy %>% 
  filter(candies == "starburst") %>% 
  filter(rating == "DESPAIR") %>% 
  summarise(despair_rating_count_starburst = n())
```
# For the next three questions, count despair as -1, joy as +1, and meh as 0.
```{r}
candy_count <- candy %>% 
  mutate(rating = str_replace_all(rating, "JOY", "1"),
         rating = str_replace_all(rating, "DESPAIR", "-1"),
         rating = str_replace_all(rating, "MEH", "0"))

candy_count <- candy_count %>% 
  mutate(rating = as.numeric(rating))
```


# 6. What was the most popular candy bar by this rating system for each gender in the dataset ?
```{r}
candy_count %>%
  filter(!is.na(rating)) %>%
  filter(gender == "Male") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_by_male = sum(rating)) %>% 
  arrange(desc(pop_candy_by_male)) %>% 
  head(1) %>% 
  ungroup()

candy_count %>%
  filter(!is.na(rating)) %>%
  filter(gender == "Female") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_by_female = sum(rating)) %>% 
  arrange(desc(pop_candy_by_female)) %>% 
  head(1) %>% 
  ungroup()

candy_count %>%
  filter(!is.na(rating)) %>%
  filter(gender == "I'd rather not say") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_by_other = sum(rating)) %>% 
  arrange(desc(pop_candy_by_other)) %>% 
  head(1) %>% 
  ungroup()

candy_count %>%
  filter(!is.na(rating)) %>%
  filter(is.na(gender)) %>% 
  group_by(candies) %>% 
  summarise(pop_candy_by_NA = sum(rating)) %>% 
  arrange(desc(pop_candy_by_NA)) %>% 
  head(1) %>% 
  ungroup()
```


# 7. What was the most popular candy bar in each year?
```{r}
candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(year == 2015) %>% 
  group_by(candies) %>% 
  summarise(pop_candy_2015 = sum(rating)) %>% 
  arrange(desc(pop_candy_2015)) %>% 
  head(1) %>% 
  ungroup()


candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(year == 2016) %>% 
  group_by(candies) %>% 
  summarise(pop_candy_2016 = sum(rating)) %>% 
  arrange(desc(pop_candy_2016)) %>% 
  head(1) %>% 
  ungroup()


candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(year == 2017) %>% 
  group_by(candies) %>% 
  summarise(pop_candy_2017 = sum(rating)) %>% 
  arrange(desc(pop_candy_2017)) %>% 
  head(1) %>% 
  ungroup()
```


# 8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
```{r}
candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(country == "USA") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_USA = sum(rating)) %>% 
  arrange(desc(pop_candy_USA)) %>% 
  head(1) %>% 
  ungroup()


candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(country == "Canada") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_Canada = sum(rating)) %>% 
  arrange(desc(pop_candy_Canada)) %>% 
  head(1) %>% 
  ungroup()


candy_count %>% 
  filter(!is.na(rating)) %>% 
  filter(country == "UK") %>% 
  group_by(candies) %>% 
  summarise(pop_candy_UK = sum(rating)) %>% 
  arrange(desc(pop_candy_UK)) %>% 
  head(1) %>% 
  ungroup()


candy_count %>% 
  filter(!is.na(rating)) %>%
  filter(!is.na(country)) %>% 
  filter(!(country %in% c("USA", "UK", "Canada"))) %>% 
  group_by(candies) %>% 
  summarise(pop_candy_others = sum(rating)) %>% 
  arrange(desc(pop_candy_others)) %>% 
  head(1) %>% 
  ungroup()
```

