---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(here)
```



```{r}
candy <- read_csv(here("clean_data/candy_clean.csv"))
```


# 1. What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
```{r}
candy %>%
  filter(!is.na(rating)) %>% 
  summarise(count_rating = n())
```

#2. What was the average age of people who are going out trick or treating?
```{r}
candy %>% 
  filter(trick_or_treating == "Yes") %>% 
  summarise(avg_age_tt = floor(mean(age)))
```

# 3. What was the average age of people who are not going trick or treating?
```{r}
candy %>% 
  filter(trick_or_treating == "No") %>% 
  summarise(avg_age_no_tt = floor(mean(age)))
```


# 4. For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
candy %>% 
  filter(rating == "JOY") %>% 
  group_by(candies, rating) %>% 
  summarise(joy_rating_per_candies = n()) %>% 
  arrange(desc(joy_rating_per_candies)) %>% 
  head(1) %>% 
  ungroup()

candy %>% 
  filter(rating == "DESPAIR") %>% 
  group_by(candies, rating) %>% 
  summarise(joy_rating_per_candies = n()) %>% 
  arrange(desc(joy_rating_per_candies)) %>% 
  head(1) %>% 
  ungroup()

candy %>% 
  filter(rating == "MEH") %>% 
  group_by(candies, rating) %>% 
  summarise(joy_rating_per_candies = n()) %>% 
  arrange(desc(joy_rating_per_candies)) %>% 
  head(1) %>% 
  ungroup()
```


# 5. How many people rated Starburst as despair?
```{r}
candy %>% 
  filter(candies == "starburst") %>% 
  filter(rating == "DESPAIR") %>% 
  summarise(despair_rating_count = n())
```

