---
title: "task_01_decathlon_data"
author: "Lloyd Lombardi"
date: "`r Sys.Date()`"
output: html_document
---
# Task 1 - Decathlon Data


## Introduction
The purpose of this Markdown file is to give an overview of Task 1 of the
**Dirty Data Project**.

The task concerned a single `.rds` file which contained decathlon data.

The data was in an **untidy** format, and therefore needed to be cleaned before
the following questions could be answered:

1. Who had the longest jump seen in the data?
2. What was the average 100m time in each competition?
3. Who had the highest total points across both competitions?
4. What was the shot-put scores for the top three competitors in each competition?
5. What was the average points for competitiors who ran the 400m in less than 50
   seconds vs. those than ran 400m in more than 50 seconds?

In the rest of this document, the following information will be presented:
- A list of assumptions made
- Steps taken to clean the data
- Answers to the analysis questions


## A list of assumptions made

- Data can be left in 'wide' format to answer analysis questions
- Events could be split into two categories
    1. Race events
    2. Athletics events
- All 'Race events' were recorded in seconds
- All 'Athletics events' were recorded in meters
- Assertive programming could be used by inputting the current World Records as
  limits
- Analysis Q4 was taken to mean, find the top 3 competitors in each competition
  and find their shot-put scores at that competition


## Steps taken to clean the data

The raw data had row names, and these needed to be changed into a column. The `rownames_to_column` function from tibble was called to perform this
```{r,eval=FALSE}
decathlon <- tibble::rownames_to_column(decathlon, var = "athlete_name")
```

The athletes' names were in an untidy format
Names were converted to title, usins `str_to_title`
```{r,eval=FALSE}
decathlon <- decathlon %>% 
  mutate(athlete_name = str_to_title(athlete_name))
```

The column names were currently untidy
Janitor `clean_names` was run
```{r,eval=FALSE}
decathlon_clean <- clean_names(decathlon)
```

To further tidy the column names, any column beginning with an 'x' was edited
so that the 'x' was removed
The `rename_with` function was employed
```{r,eval=FALSE}
decathlon_clean <- decathlon_clean %>% 
  rename_with(~ gsub("^x", "", .x))
```

To make the data frame easier to read, the rank, points and competition columns
were moved to after the athletes' names
This was completed by using `relocate`
```{r,eval=FALSE}
decathlon_clean <- decathlon_clean %>% 
  relocate(c(rank:competition), .after = athlete_name)
```



The first vector was used to add 'seconds' as units to the columns of races
```{r,eval=FALSE}
decathlon_clean <- decathlon_clean %>% 
  rename_with(~str_replace(., "$", "_(s)"), .cols = starts_with(race_columns))
```

The second vector was used to add 'meters' as units to the columns of athletics
```{r,eval=FALSE}
decathlon_clean <- decathlon_clean %>%
  rename_with(~str_replace(., "$", "_(m)"), .cols = starts_with(athletics_columns))
```

Some assertive programming was coded to verify the various scores
The current World Records for each individual event was used for the limits
```{r,eval=FALSE}
decathlon_clean %>% 
  verify(`100m_(s)` > 9.85) %>% 
  verify(`long_jump_(m)` < 8.95) %>% 
  verify(`shot_put_(m)` < 23.37) %>% 
  verify(`high_jump_(m)` < 2.45) %>% 
  verify(`400m_(s)` > 43.03) %>% 
  verify(`110m_hurdle_(s)` > 12.8) %>% 
  verify(`discus_(m)` < 74.08) %>% 
  verify(`pole_vault_(m)` < 6.21) %>% 
  verify(`javeline_(m)` < 98.48) %>% 
  verify(`1500m_(s)` > 230)
```


## Answers to the analysis questions

1. Who had the longest jump seen in the data?
```{r,eval=FALSE}
Clay - 7.96m
```

2. What was the average 100m time in each competition?
```{r,eval=FALSE}
Decastar - 11.18s
OlympicG - 10.92s
```

3. Who had the highest total points across both competitions?
```{r,eval=FALSE}
Sebrle - 17110 points
```

4. What was the shot-put scores for the top three competitors in each competition?
```{r,eval=FALSE}
Decaster:
              1. Sebrle   - 14.83m
              2. Clay     - 14.26m
              3. Karpov   - 14.77m
OlympicG:
              1. Sebrle   - 16.36m
              2. Clay     - 15.23m
              3. Karpov   - 15.93m
```

5. What was the average points for competitiors who ran the 400m in less than 50
   seconds vs. those than ran 400m in more than 50 seconds?
```{r,eval=FALSE}
Average points for competitors who ran 400m in less than 50s - 8120 points
Average points for competitors who ran 400m in more than 50s - 7727 points
```


You can access the **Github Repo** [here](https://github.com/lloydlombardi/dirty_data_codeclan_project_lloyd_lombardi/tree/main/task_01_decathlon_data) 